---
title: "A Homework of Biblical Proportions"
author: "Sanath Shivaswamy & Christian Ghorbani"
date: "April 29, 2016"
output: html_document
---

In this HW, which I borrowed from Professor Mahmoud Parsian's course on Big Data (highly recommended), we will analyze the text of the bible. The ascii version resides in the file "ascii_bible.txt" on Camino. Use an editor to familiarize yourself with the structure of the file. Then perform the following operations, listed below as questions for you to answer. 

## Q1: Read in the file using any R function you like and store each verse in a text array. After which print the top 20 verses. (Remove the top two lines which contain the title.)


```{r}
library(stringr)
bible_full = readLines("ascii_bible.txt")
bible_text = str_replace(bible_full,"^\\s+","") #remove leading whitespaces 
res = grep("[0-9][0-9][0-9]:[0-9][0-9][0-9]",bible_text)
temp = NULL
for (i in seq(1:(length(res)-1))){
  x = paste(bible_text[res[i]:(res[i+1]-1)],collapse = " ")
  temp = c(temp,x)
}
verses = c(temp,paste(bible_text[res[length(res)]:length(bible_text)],collapse = " "))
verses[1:20]
```

## Q2: How many verses are there in total? 

```{r}
length(verses)
# 31102 verses in bible
```

## Q3: Each verse has the number "CCC:VVV" where CCC is the chapter number and VVV is the verse number. How many chapters are there? 

```{r}
chap = grep("[0-9][0-9][0-9]:001",verses)
length(chap)
# 1189 Chapters in bible
```

## Q4: Extract an array for the verse numbers, and also one for the verse text.

```{r}
ver_no = c(substr(verses,5,7)) #array of verse numbers
ver_txt =  str_split_fixed(verses, "[0-9][0-9][0-9]:[0-9][0-9][0-9] ",2)
ver_txt = ver_txt[,2]
head(ver_no)
head(ver_txt)
```

## Q5: Lower case all text.

```{r}
ver_txt = str_to_lower(ver_txt)
```


## Q6: Convert the text of all verses into a Corpus using the **tm** package. 

```{r}
library(tm)
ver_txt = Corpus(VectorSource(ver_txt))
ver_txt
```

## Q7: Remove all punctuation. Use a corpus function for this. How many unique words are there? 

```{r}
ver_txt = tm_map(ver_txt,removePunctuation)
ver_uniq = DocumentTermMatrix(ver_txt)
ver_uniq 
#12651 unique words
```

## Q8: Remove all stopwords. Now how many unique terms are there? 

```{r}
ver_stp = ver_txt
ver_stp = tm_map(ver_stp,removeWords,stopwords("english"))
ver_uniq = DocumentTermMatrix(ver_stp)
ver_uniq 
#12555 unqiue words
```

## Q9: Now stem the text, and see how many unique words there are.

```{r}
library(SnowballC)
ver_stem = tm_map(ver_stp,stemDocument)
ver_uniq = DocumentTermMatrix(ver_stem)
ver_uniq
#9125 unique words
```

## Q10: Convert the corpus into a term-document matrix (TDM), where each verse is a separate document. How many distinct words are there in the bible?

```{r}
ver_tdm = DocumentTermMatrix(ver_stem)
ver_tdm
#9125 distinct words in the bible
```

## Q11: Convert the TDM into a matrix and find the 50 most common words in the bible. 

```{r}
spr_rmv = removeSparseTerms(ver_tdm,0.99) 
ver_mat = as.matrix(spr_rmv)
cm_words = sort(colSums(ver_mat), decreasing = TRUE) #frequency for each term
w_50 = head(cm_words,50)
w_50 #50 most common words
```

## Q12: Make a wordcloud of the top 100 words in the bible. 

```{r}  
library(wordcloud)  
w_100 = head(cm_words,100) 
w_100_nms = names(w_100) #names of top 100 words
wordcloud(w_100_nms,w_100) #creating word cloud
```

## Q13: Mood score the original text of the bible (before stemming)

```{r}
#creating positive and negitive words
HIDict = readLines("inqdict.txt")
dict_pos = HIDict[grep("Pos",HIDict)]
poswords = NULL
for (s in dict_pos) {
    s = strsplit(s,"#")[[1]][1]
    poswords = c(poswords,strsplit(s," ")[[1]][1])
}
dict_neg = HIDict[grep("Neg",HIDict)]
negwords = NULL
for (s in dict_neg) {
    s = strsplit(s,"#")[[1]][1]
    negwords = c(negwords,strsplit(s," ")[[1]][1])
}
poswords = tolower(poswords)
negwords = tolower(negwords)
poswords = unique(poswords)
negwords = unique(negwords)

#Mood Scoring the bible text
ver_txt =  str_split_fixed(verses, "[0-9][0-9][0-9]:[0-9][0-9][0-9] ",2)
ver_txt = ver_txt[,2] #bible text
ver_txt = Corpus(VectorSource(ver_txt))
ver_txt = tm_map(ver_txt,removePunctuation)
v1 = data.frame(text = sapply(ver_txt, as.character), stringsAsFactors = FALSE) #corpus to dataframe
v1 = str_replace(v1$text,"^\\s+","")  #remove leading whitespaces
v1 = trimws(v1,which = "right") # triming trailing whitespaces
v1 = str_to_lower(v1) 
ver_txt1 = unlist(strsplit(v1," ")) #spliting by space to get words
posmatch = match(ver_txt1,poswords)
numposmatch = length(posmatch[which(posmatch>0)])
negmatch = match(ver_txt1,negwords)
numnegmatch = length(negmatch[which(negmatch>0)])
print(c(numposmatch,numnegmatch))
#No of Positive words = 56827
#No of Negative words = 38769
```



## Q14: Summarize the bible into 100 verses. (Or some fraction of the total number of verses, it's your choice.) Be super careful here as this may take a long time unless you are clever about it, or find some elegant way to speed things up!

```{r}
a = c(2,2)
b = c(2,2)
x = data.frame(a,b)
y = data.frame(a,b)
z = data.frame(a,b)
column_names = c("Pos_Score","Verse")
x = `colnames<-`(x,column_names)
counter = 1
for(verse in v1){
  ver_txt1 = unlist(strsplit(verse," "))
  posmatch = match(ver_txt1,poswords)
  numposmatch = length(posmatch[which(posmatch>0)])
  if(numposmatch > 5){ #taking only those verses which have positive mood score > 5
    x[counter,1] = numposmatch
    x[counter,2] = verse
    counter = counter + 1
  }
}

sum_txt = x$Verse #Taking only verses

#Function to generate Summary
text_summary = function(text, n) {
  m = length(text)  # No of sentences in input
  jaccard = matrix(0,m,m)  #Store match index
  for (i in 1:m) {
    for (j in i:m) {
      a = text[i]; aa = unlist(strsplit(a," "))
      b = text[j]; bb = unlist(strsplit(b," "))
      jaccard[i,j] = length(intersect(aa,bb))/
                          length(union(aa,bb))
      jaccard[j,i] = jaccard[i,j]
    }
  }
  similarity_score = rowSums(jaccard)
  res = sort(similarity_score, index.return=TRUE,
          decreasing=TRUE)
  idx = res$ix[1:n]
  summary = text[idx]
}

#Summary of bible with 100 verses
res = text_summary(sum_txt,100)
res

```


```{r}
#Working code to produce a Negative Summary
#a = c(2,2)
#b = c(2,2)
#x = data.frame(a,b)
#y = data.frame(a,b)
#z = data.frame(a,b)
#column_names = c("Neg_Score","Verse")
#x = `colnames<-`(x,column_names)
#counter = 1
#for(verse in v1){
#  ver_txt1 = unlist(strsplit(verse," "))
#  negmatch = match(ver_txt1,negwords)
#  numnegmatch = length(negmatch[which(negmatch>0)])
#  if(numnegmatch > 5){ #taking only those verses which have positive mood score > 5
#    x[counter,1] = numnegmatch
#    x[counter,2] = verse
#    counter = counter + 1
#  }
#}

#sum_txt = x$Verse

#res = text_summary(sum_txt,100)
#head(res)
```


```{r}
#Working code to produce a Mix of Positive and Negative Summary
#a = c(2,2)
#b = c(2,2)
#x = data.frame(a,b)
#y = data.frame(a,b)
#z = data.frame(a,b)
#column_names = c("Verse")
#x = `colnames<-`(x,column_names)
#counter = 1
#for(verse in v1){
#  ver_txt1 = unlist(strsplit(verse," "))
#  posmatch = match(ver_txt1,poswords)
#  numposmatch = length(posmatch[which(posmatch>0)])
#  negmatch = match(ver_txt1,negwords)
#  numnegmatch = length(negmatch[which(negmatch>0)])
#  if(numposmatch > 5 | numnegmatch > 5){ #taking only those verses which have positive mood score > 5
#    x[counter,1] = verse
#    counter = counter + 1
#  }
#}
#sum_txt = x$Verse

#Summary of bible with 100 verses
#res = text_summary(sum_txt,100)
#head(res)
```

